\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-16 popl-2019-journal-v/paper.tex
Fix a lot of R1 remarks.

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}

@@

% prevent a compilation failure with JFP.
\undef{\base}
\input{../fully-abstract-comp/preamble}
% Paper specific redefinitions of commands
\renewcommand{\MemFrag}{\shareddom{MemFrag}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\encInstr}[1]{\plainfun{encode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}
\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
\renewcommand{\comp}{\var{comp}}
\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}

@@

\begin{jversion}
\subsection{Decoding and encoding functions}
The operational semantics of the capability machine uses the function $\decInstr{}$ to decode instructions.
We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
The $\decInstr{} : \Word \fun \Instr$ should be surjective and injective for all non-$\fail$ instructions.
Further, it should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
\[
\forall c \in \Caps \ldotp \decInstr{c} = \fail
\]
The $\encInstr{} : \Instr \fun \ints$ function should be injective, and  $\decInstr{}$ is its left inverse, i.e.
\[
\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
\]
These assumptions are sufficient to construct program examples in terms of instructions rather than machine words (using $\encInstr{}$), and run them on the machine (using the fact that $\decInstr{}$ is the left inverse of $\encInstr{}$).
% The $\encInstr{}$ function allows us to specify the program ``abstractly''.
% As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
%This is not to say that encoding is an easy problem. There are many things to get.

The machine also assumes decode and encode functions for permissions $\decPerm{} : \Perm \fun \ints$ and $\encPerm{} : \ints \fun \Perm$.
We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective. For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the $\tgetp{}{}$ error value -1, i.e.
\[
  \forall \perm \in \Perm \ldotp \encPerm{\perm} \neq -1
\]
For linearity we assume similar functions.
Finally in the interpretation of $\tisptr{}{}$, the machine uses an encode function for word types $\encType{}$.
This function encodes each kind of word as an integer.
This is very much like the previous functions.
It encodes each kind of word differently and all words of the same kind to the same integer.
\end{jversion}

\subsection{Components, linking, programs, and contexts }
\label{subsec:components-linking}
The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
To clarify this, we introduce notions of components and programs from which we construct executable configurations.
A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
It has exports that can satisfy the imports of other components.
% Structure
A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
An export associates a word with a symbol.
%base = library
A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
The latter can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together, as long as only one has a main function.
Two components can be linked when their memories, seals, and linear addresses are disjoint.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.

@@

% known stack base
% + more
At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
However, this still does not ensure that the caller's stack frame is on top of the call stack.
The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.

@@

\begin{figure}[htb]
  \centering
{\small
\[
  \begin{array}{r >{$}p{5.1cm}<{$} r l}
\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
      &    &26: & \tmove{\rtmp{2}}{0}\\
  \end{array}
\]
}
\caption{
  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
  % Variable $\offpc$ is the offset from the first instruction of the call to the set-of-seals, and $\offsigma$ is the offset to the seal within the set.
  % Instructions 1-3: ensure that the stack is non-empty.
  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
  % Instructions 6-9: Load the seal designated for this call.
  % Instruction 10: Seal capability for private stack to create data return pointer.
  % Instructions 11-13: Seal capability for return point to create code return pointer.
  % Instructions 14-15: Clean up temporary register and jump.
  % Instructions 16-26: The return code.
  % Instructions 16-23: check the stack base.
  % Instruction 24: join the stack pointer.
  % Instructions 25-26: pop 42 from the stack.
  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set-of-seals it uses and $\offsigma$ the offset in the set-of-seals to the call seal.
  $\stkb$ is the globally agreed on stack base.
  There are some magic numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
  Line 7: $-5$, offset from line 6 (where $\pcreg$ was copied into $\rtmp{1}$) to line 1.
  Line 12: $5$, offset to the return address.
  Line 19: $5$, offset to fail.
  Line 21: offset to address after fail.
}
  \label{fig:call-code}
\end{figure}

@@

    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and

@@

    \begin{multlined}

@@

  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data and set-of-seals capabilities.
  The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
  This makes a clear separation between code and data memory as neither can contain capabilities for the other.
  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking return seals that were indirectly accessible through the data structure.
  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most have read-execute permission and all capabilities for data memory at most have read-write permission.
  A component's exports should be sealed code or data capabilities, sealed with a closure seal.

@@

\begin{figure}
\begin{mathpar}
  \inferrule*[Right=Base]{
    \dom(\mscode) = [\baddr,\eaddr]\\
    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
    \dom(\msdata) \mathrel{\#} \ta \\
    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
  }{
    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
  }
  \and
  \inferrule*[Right=Main]{
    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
    \ta \vdash \var{comp}_0 \\
    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
  }{
    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
  }
\end{mathpar}
\caption{Well-formedness judgement.}
\label{fig:well-formed}
\end{figure}

@@

The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
Further, the base component should satisfy the well-formedness judgement.
The \textsc{Base} rule ensures that the component has a certain structure.
Roughly speaking, the \textsc{Base} rule requires the following:
\begin{itemize}
\item \textit{Code and data memory are disjoint.}
\item \textit{Code memory is padded with zeros ($\mspad$) and this padded memory may not be referenced.}
  This prevents code memories from being spliced together.
  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
  Further, the possible control-flows of a component suddenly depend on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
\item \textit{All memory can either be addressed by any number of non-linear capabilities or at most one linear capability.}
  The data address space is split into $A_\var{non-linear}$ and $A_\var{linear}$ that can be addressed by non-linear capabilities and linear capabilities, respectively.
  The judgement ensure uniqueness of linear capabilities by letting each address of the data memory take sole ownership of addresses in $A_\var{linear}$.
\item \textit{All import addresses are part of the data memory.} 
\item \textit{Import and export symbols are disjoint.}
\item \textit{One of the following is true}
  \begin{itemize}
  \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code. We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to closure seals that could be used to protect calls.}.
  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code. We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
  \end{itemize}
\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
\item \textit{The code memory satisfies the component-code well-formedness judgement (\figurename~\ref{fig:well-formed-code})}, allowing only seal capabilities for seals owned by the component and numbers. If the numbers correspond to the encoding of a call instruction, then they must reference a unique return seal in the code memory.
\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address. It allows only numbers, linear and non-linear capabilities to memory owned by the component and values sealed with closure seals.
\item \textit{All the exports satisfy the components-export well-formedness judgement (\figurename~\ref{fig:well-formed-export})}, requiring that they are well-formed by the component-word judgement and don't own linear memory.
\end{itemize}

@@

@@

The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that applies to unsafe source languages.
Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.

>>>>

\usepackage[utf8]{inputenc}
% \usepackage{amsmath}
% \usepackage{amssymb}

@@

% prevent a compilation failure with JFP.
\undef{\base}
\input{../fully-abstract-comp/preamble}
% Paper specific redefinitions of commands
\newcommand{\MemFrag}{\shareddom{MemFrag}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\encInstr}[1]{\plainfun{encode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}
\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
\renewcommand{\comp}{\var{comp}}
\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}

@@

\begin{jversion}
\subsection{Decoding and encoding functions}
The operational semantics of the capability machine uses the function $\decInstr{}$ to decode instructions.
We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
The $\decInstr{} : \Word \fun \Instr$ should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
\[
\forall c \in \Caps \ldotp \decInstr{c} = \fail
\]
The $\decInstr{}$ should have a right inverse $\encInstr{} : \Instr \fun \ints$, i.e.
\[
\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
\]
These assumptions are sufficient to construct program examples in terms of instructions rather than machine words (using $\encInstr{}$), and run them on the machine (using the fact that $\decInstr{}$ is the left inverse of $\encInstr{}$).
% The $\encInstr{}$ function allows us to specify the program ``abstractly''.
% As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
%This is not to say that encoding is an easy problem. There are many things to get.

The machine also assumes decode and encode functions for permissions $\encPerm{} : \Perm \fun \ints$ and $\decPerm{} : \ints \fun \Perm$.
We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$.
For $\encPerm{}$, we assume that it does not encode anything to the $\tgetp{}{}$ error value -1, i.e.
\[
  \forall \perm \in \Perm \ldotp \encPerm{\perm} \neq -1
\]
For linearity we assume similar functions.
Finally in the interpretation of $\tisptr{}{}$, the machine uses an encode function for word types $\encType{}$.
This function encodes each kind of word as an integer.
This is very much like the previous functions.
It encodes each kind of word differently and all words of the same kind to the same integer.
\end{jversion}

\subsection{Components, linking, programs, and contexts}
\label{subsec:components-linking}
The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
To clarify this, we introduce notions of components and programs from which we construct executable configurations.
A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
It has exports that can satisfy the imports of other components.
% Structure
A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
An export associates a word with a symbol.
%base = library
A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
The latter can be seen as a program that still needs to be linked with libraries.
% Linking .
Components are combined into new components by linking them together, as long as only one has a main function.
Two components can be linked when their memories, seals, and linear addresses are disjoint.
They are combined by taking the union of each of their constituents.
For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.

@@

% known stack base
% + more
At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
However, this still does not ensure that the caller's stack frame is on top of the call stack.
The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
An adversarial callee may intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
At this point (Figure~\ref{fig:stack-base-abuse-d}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.

@@

\begin{figure}[htb]
  \centering
{\small
\[
  \begin{array}{r >{$}p{5.1cm}<{$} r l}
\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \tcca{\rtmp{2}}{1}\\
9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
      &    &26: & \tmove{\rtmp{2}}{0}\\
  \end{array}
\]
}
\caption{
  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
  % Variable $\offpc$ is the offset from the first instruction of the call to the set-of-seals, and $\offsigma$ is the offset to the seal within the set.
  % Instructions 1-3: ensure that the stack is non-empty.
  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
  % Instructions 6-9: Load the seal designated for this call.
  % Instruction 10: Seal capability for private stack to create data return pointer.
  % Instructions 11-13: Seal capability for return point to create code return pointer.
  % Instructions 14-15: Clean up temporary register and jump.
  % Instructions 16-26: The return code.
  % Instructions 16-23: check the stack base.
  % Instruction 24: join the stack pointer.
  % Instructions 25-26: pop 42 from the stack.
  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
  The number $\offpc$ is the offset between the memory location where the first instruction of the call is stored to the memory location where the set-of-seals capability is stored that should be used for sealing the return capability.
  This set-of-seals capability may give access to a range of seals and the offset $\offsigma$ identifies which of these should be used.
  $\stkb$ is the globally agreed on stack base.
  There are some magic numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
  Line 7: $-5$, offset from line 6 (where $\pcreg$ was copied into $\rtmp{1}$) to line 1.
  Line 12: $5$, offset to the return address.
  Line 19: $5$, offset to fail.
  Line 21: offset to address after fail.
}
  \label{fig:call-code}
\end{figure}

@@

    $\opc, c_\opc =\var{setupOpc}(\Phi.\reg(\pcreg))$ and

@@

    \begin{multlined}[t]

@@

  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data (including encoded instructions) and set-of-seals capabilities.
  The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
  This makes a clear separation between code and data memory as neither can contain capabilities for the other.
  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking return seals that were indirectly accessible through the data structure.
  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most have read-execute permission and all capabilities for data memory at most have read-write permission.
  A component's exports should be sealed code or data capabilities, sealed with a closure seal.

@@

\begin{figure}
\begin{mathpar}
  \inferrule*[Right=Base]{
    \dom(\mscode) = [\baddr,\eaddr]\\
    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
    \dom(\msdata) \mathrel{\#} \ta \\
    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
    \forall w_{\mathrm{export}} \in \overline{w_{\mathrm{export}}} \ldotp \dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}
  }{
    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
  }
  \and
  \inferrule*[Right=Main]{
    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
    \ta \vdash \var{comp}_0 \\
    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
  }{
    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
  }
\end{mathpar}
\caption{Well-formedness judgement.}
\label{fig:well-formed}
\end{figure}

@@

The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
The \textsc{Main} rule requires the component to have an entry point in the form of a pair of capabilities $c_{\mathrm{main},c}, c_{\mathrm{main},d}$, which should be part of the exports.
Further, the base component should satisfy the well-formedness judgement.
The \textsc{Base} rule ensures that the component has a certain structure.
Roughly speaking, the \textsc{Base} rule requires the following:
\begin{itemize}
\item \textit{Code and data memory are disjoint.}
\item \textit{Code memory is padded with zeros ($\mspad$) and this padded memory may not be referenced.}
  This prevents code memories from being spliced together.
  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
  Further, the possible control-flows of a component suddenly depend on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
\item \textit{All memory can either be addressed by any number of non-linear capabilities or at most one linear capability.}
  The data address space is split into $A_\var{non-linear}$ and $A_\var{linear}$ that can be addressed by non-linear capabilities and linear capabilities, respectively.
  The judgement ensure uniqueness of linear capabilities by letting each address of the data memory take sole ownership of addresses in $A_\var{linear}$.
\item \textit{All import addresses are part of the data memory.} 
\item \textit{Import and export symbols are disjoint.}
\item \textit{One of the following is true}
  \begin{itemize}
  \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code. We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to closure seals that could be used to protect calls.}.
  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code. We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
  \end{itemize}
\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
\item \textit{The code memory satisfies the component-code well-formedness judgement (\figurename~\ref{fig:well-formed-code})}, allowing only seal capabilities for seals owned by the component and numbers. If the numbers correspond to the encoding of a call instruction, then they must reference a unique return seal in the code memory.
\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address. It allows only numbers, linear and non-linear capabilities to memory owned by the component and values sealed with closure seals.
\item \textit{All the exports satisfy the components-export well-formedness judgement (\figurename~\ref{fig:well-formed-export})}, requiring that they are well-formed by the component-word judgement and don't own linear memory.
\end{itemize}

@@

      \item There exist $\perm$, $\lin$, $\baddr$, $\eaddr$, $\aaddr$, $\aaddr'$ such that all of the following hold:

@@

The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that applies to unsafe source languages.
Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.

In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.
Our full abstraction result does not apply to those scenarios because it is intended to be used in the verification of a secure compiler where such scenarios are not relevant.
Nevertheless, we believe our Fundamental Theorem of Logical Relations (see Section~\ref{thm:ftlr}) does apply in such scenarios because it is step-bounded.
To deal with the evolving classification of components as trusted or adversarial, we can simply apply the theorem several times with different choices of $\ta$ and taking $n$ small enough that the compromise of $\ta$ has not yet happened until that point. 
Because the theorem only requires reasonability up to $n$ steps, we get well-bracketedness guarantees up to the right step in the execution.
This can be seen as a step-bounded version of the idea explained in \citep{patrignani_modular_2016}.
It is an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenarios without the need to build in dynamic compromise scenarios into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
